AtmoSwing (v1.4): Analog Technique Model for Statistical weather forecastING

### 2015-10-21 AtmoSwing - Geoscientific Model Development/template.tex
Addition of the discussion on the analogue method.

The evolution of the analogue method for precipitation forecasting led to several improvements. However, some of these are specific for a certain region and are not relevant for others, and some perform better depending on the lead time. Thereby, we cannot provide a single best parameterization for the method, neither rank them from worst to best. We will thus provide some examples of parameterizations of the most recent method variant, in the chronological order they appeared. The most common usage of the analogue method is for precipitation forecasting. Thus, the evolution presented hereafter focuses on this predictand.

@@

The use of the analogue method for operational forecasting of daily precipitation originates in the work of \citet{Duband1970,Duband1974,Duband1981}. It was then designed for operational forecasting at EDF (\'{E}lectricit\'{e} de France) in order to better manage water resources and flood risks. The geopotential height at 700~hPa and 00~h was used for its higher stability than the sea level pressure (SLP) field, and its sensitivity to the atmospheric disturbances \citep{Guilbaud1997}. However, the SLP at 06~h was also used for determining the intensity of precipitation, as well as the temperature at 700~hPa as indicator of the thermal state of the air mass \citep{Duband1974}. The data were then based on 37 radiosounding over Europe and condensed by a principal component analysis. The first 25 analogues on the geopotential at 700~hPa were considered, and multiple local regressions between precipitation, SLP, and the temperature at 700~hPa were established on these analogues before being applied to the target situation \citep{Guilbaud1997}. The archives were split by season to consider comparable situations in terms of distribution of solar energy \citep{Lorenz1969}. This initial rigid division into seasons was then transformed into a moving selection of more or less two months around the target date. The forecast was then performed only on the basis of observations and was temporally extrapolated to the following two days. After a few years, this statistical temporal extrapolation was abandoned in favor of a statistical adaptation using predictors resulting from an global numerical model, allowing a forecast over the following 4~days.

Thereafter, the geopotential was considered at 1000~hPa and 700~hPa and the analogues selection was done in two stages: first according to an Euclidean distance in the space of the principal components of the geopotential field at 700~hPa, and then according to a correlation criterion (on the geopotential field at 700 and 1000~hPa, as well as on the thickness between these two levels) in order to remove days too dissimilar \citep{Guilbaud1997}. The regression approach to determining precipitation was eventually abandoned in favor of the probabilistic forecasting (synthesized by the percentiles 20, 60 and 90~\%), as it is performed today by interpolating linearly over the cumulative empirical distribution of daily precipitation measured at the analogue days \citep{Guilbaud1997,Guilbaud1998}.

@@

@@

\subsection{Discussion}
% TODO: write

%\citet{Bliefernicht2010} a observé que la performance de la méthode est plus élevée en hiver qu'en été. Ainsi, la relation entre les prédicteurs synoptiques et les précipitations locales est plus faible en été, en raison des pluies d'origine convective qui ont un caractère plus aléatoire.

%La méthode des analogues, ou une forme proche, a également été utilisée pour la prévision à court ou moyen terme des températures journalières \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, des vents \citep{Gordon1987}, des avalanches \citep{Obled1980, Bolognesi1993}, de l'insolation \citep{Bois1981}, et de la trajectoire de cyclones tropicaux \citep{Keenan1981, Sievers2000, Fraedrich2003}.

%\citet{Guilbaud1997} a effectué une recherche bibliographique des utilisations de la méthode en prévision à long terme et a identifié des applications opérationnelles pour des prévisions mensuelles dans de nombreux pays, notamment au Canada \citep{Shabbar1986}, en Hongrie \citep{Toth1989}, aux Pays-Bas \citep{Nap1981}, et en Angleterre \citep{Murray1974}, ainsi que des prévisions saisonnières : \citet{Barnett1978}, \citet{Bergen1982} et \citet{Livezey1988}.

%L'utilisation de la méthode des analogues en tant que technique d'adaptation de modèles climatiques fait l'objet d'un nombre croissant d'études \citep{Zorita1999,Wetterhall2005,Matulla2008}. \citet{Wetterhall2007} ont également utilisé la méthode des analogues en Suède, dans le cadre d'une étude sur le réchauffement climatique, et la trouvent simple, mais moins performante que d'autres méthodes plus complexes. Cependant, ils n'utilisent que la pression au niveau de la mer en tant que prédicteur, et considèrent uniquement la meilleure analogue. Ces deux aspects ne sont clairement pas optimaux, et une qualité médiocre de la méthode n'est alors pas surprenante. L'implémentation actuelle de la méthode (voir section \ref{sec:etat_art:implementation_actuelle}) fournit clairement des résultats plus pertinents.

%Une version de la méthode a été évaluée dans le projet STARDEX \citep[\textit{STAtistical and Regional dynamical Downscaling of EXtremes for European regions}, voir][]{Goodess2003, Stardex2005}. Un des objectifs du projet étant de comparer différentes méthodes de descente d'échelle pour la détermination des extrêmes météorologiques, la méthode des analogues a été mise en concurrence avec d'autres techniques et a été sélectionnée parmi les plus intéressantes \citep{Maheras2005, Schmidli2007}.

%\citet{GibergansBaguena2007} ont implémenté la méthode pour des prévisions de précipitations journalières en Catalogne. Ils ont considéré le géopotentiel à 700 et 1000~hPa, ainsi que l'épaisseur entre ces deux niveaux pour la recherche d'analogie de circulation, puis ont appliqué une seconde sélection sur la base de données locales obtenues par radiosondage: indices de stabilité, eau précipitable, température, etc...

%\citet{Hamill2006} ont utilisé une approche par analogie sur les reprévisions du modèle global des États-Unis (GFS, voir sections \ref{sec:donnees:previsions_GFS} et \ref{sec:donnees:resimulations_GFS}). Cette application a permis de corriger des erreurs systématiques dans les prévisions d'ensemble des températures et des précipitations. En effet, l'approche statistique a mis en évidence certains biais dans les prévisions du modèle numérique, qui ont pu être ajustés par la prise en compte de la climatologie locale intrinsèque à la méthode des analogues. De plus, la dispersion des analogues étant plus importante que celle des prévisions d'ensemble, la sous-dispersion de ces dernières a pu être corrigée à l'aide des analogues \citep{Hamill2006}. Une des raisons invoquées pour la bonne performance de la méthode des analogues par rapport au modèle numérique global est le fait que ce dernier a une orographie lissée, qui ne tient pas compte des réelles spécificités locales, ce que permet l'approche par analogie.

%La méthode présente cependant des limites dont il faut être conscient. La première est la nécessité de disposer d'une archive conséquente de la variable cible à un pas de temps pertinent, ici les précipitations. Sans une base de données de prédictands de plusieurs décennies, la méthode n'est pas applicable.

%Les archives des prédicteurs sont également importantes et plus particulièrement leur homogénéité. Nous disposons actuellement de jeux de données de réanalyses (voir section \ref{sec:donnees:reanalyses1_ncep}) de bonne qualité, mais contenant tout de même des discontinuités significatives dues à la densification des réseaux de mesures et au déploiement de nouvelles techniques de mesure et de l'imagerie satellitaire. Nous ne pourrons a priori jamais complètement éliminer cette hétérogénéité, car les données demeureront manquantes. Toutefois, la qualité de ces données les rend tout à fait utilisables pour la méthode des analogues, et ceci partout sur le globe !

%Un aspect supplémentaire concerne les différences entre les prédicteurs disponibles en temps réel et les archives de ces prédicteurs. En effet, les deux jeux de données n'ont généralement pas la même climatologie, car ils ne résultent pas du même modèle, et n'ont pas été élaborés de la même manière. Il existe des alternatives, telles que l'utilisation en opérationnel de prévisions émises par un modèle figé, conjointement à des archives élaborées par ce même modèle en reprévision.

%Toutefois, les prévisions que nous utiliserions de cette manière ne seraient pas à l'état de l'art, et les reprévisions (archives) contiendraient des erreurs et des incertitudes beaucoup plus importantes que des réanalyses. Il convient donc d'utiliser des prédicteurs robustes qui dépendent au minimum des spécificités du modèle numérique.

%Une autre limite réside dans le fait que les événements extrêmes peuvent être sous-représentés dans l'échantillon d'analogues considéré. En effet, dans une archive météorologique d'une durée limitée, les événements correspondants à des temps de retour élevés ne sont pas très nombreux. Leur représentation est certainement inférieure au nombre d'analogues que nous retenons, ce qui peut biaiser la prévision.

%Enfin, \citet{Bliefernicht2007} ont montré que les prévisions sont moins bonnes en été. Ceci est dû à la part plus importante de la convection durant les mois chauds. Les variables de la circulation synoptique ne sont en effet pas à même de prévoir la localisation des cellules orageuses. Ceci a également été constaté par \citet{BenDaoud2010}, qui avait mis en place un modèle spécifique pour les mois d'été (du 15 juin au 15 septembre). Nous pouvons estimer que les prédicteurs sont à reconsidérer (variables, niveaux, etc.), et que leur domaine peut être différent. Le problème est que, durant cette période, on peut également rencontrer situations classiques avec une circulation atmosphérique dominante comme dans les autres saisons, ou des situations de marais barométriques accompagnées d'un fort risque de convection locale, sans signature dans les variables synoptiques.

>>>>

The evolution of the analogue method for precipitation forecasting led to several improvements. However, some of these are specific for a certain region and are not relevant for others, and some perform better depending on the lead time. Thereby, we cannot provide a single best parameterization for the method, neither rank them from worst to best. We will thus provide some examples of parameterizations of the most recent method variant, in the chronological order they appeared. The most common usage of the analogue method is for precipitation forecasting. Thus, the evolution presented hereafter focuses on this predictand. However, the analogue method, or an equivalent, was also used for short to medium term forecasting of daily temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, wind \citep{Gordon1987}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}. \citet{Guilbaud1997} performed a literature review about the use of the analogue method in long-term forecasting and identified operational applications for monthly forecasts in many countries, including Canada \citep{Shabbar1986},  Hungary \citep{Toth1989}, the Netherlands \citep{Nap1981}, and England \citep{Murray1974}, as well as seasonal forecasts: \citet{Barnett1978}, \citet{Bergen1982} and \citet{Livezey1988}.

@@

The use of the analogue method for operational forecasting of daily precipitation originates in the work of \citet{Duband1970, Duband1974, Duband1981}. It was then designed for operational forecasting at EDF (\'{E}lectricit\'{e} de France) in order to better manage water resources and flood risks. The geopotential height at 700~hPa and 00~h was used for its higher stability than the sea level pressure (SLP) field, and its sensitivity to the atmospheric disturbances \citep{Guilbaud1997}. However, the SLP at 06~h was also used for determining the intensity of precipitation, as well as the temperature at 700~hPa as indicator of the thermal state of the air mass \citep{Duband1974}. The data were then based on 37 radiosounding over Europe and condensed by a principal component analysis. The first 25 analogues on the geopotential at 700~hPa were considered, and multiple local regressions between precipitation, SLP, and the temperature at 700~hPa were established on these analogues before being applied to the target situation \citep{Guilbaud1997}. The archives were split by season to consider comparable situations in terms of distribution of solar energy \citep{Lorenz1969}. This initial rigid division into seasons was then transformed into a moving selection of more or less two months around the target date. The forecast was then performed only on the basis of observations and was temporally extrapolated to the following two days. After a few years, this statistical temporal extrapolation was abandoned in favor of a statistical adaptation using predictors resulting from an global numerical model, allowing a forecast over the following 4~days.

Thereafter, the geopotential was considered at 1000~hPa and 700~hPa and the analogues selection was done in two stages: first according to an Euclidean distance in the space of the principal components of the geopotential field at 700~hPa, and then according to a correlation criterion (on the geopotential field at 700 and 1000~hPa, as well as on the thickness between these two levels) in order to remove days too dissimilar \citep{Guilbaud1997}. The regression approach to determining precipitation was eventually abandoned in favor of the probabilistic forecasting (synthesized by the percentiles 20, 60 and 90~\%), as it is performed today by interpolating linearly over the cumulative empirical distribution of daily precipitation measured at the analogue days \citep{Guilbaud1997, Guilbaud1998}.

@@

\citet{Gibergans-Baguena2007} implemented the method for daily precipitation forecasts in Catalonia. They considered the geopotential at 700 and 1000~hPa, and the thickness between these two levels, to search for analogy of circulation, and then applied a second selection on local data from radiosonde: stability indexes, precipitable water, temperature, etc.

@@

\subsection{Discussion on the method}

A version of the analogue method was evaluated during the project STARDEX \citep[\textit{STAtistical and Regional dynamical Downscaling of EXtremes for European regions}, see][]{Goodess2003, Stardex2005}. One of the project goals was to compare various downscaling methods for the determination of weather extremes, and the analogue method was selected among the most interesting from various techniques \citep{Maheras2005, Schmidli2007}. Its use as adaptation method of climate models is also the subject of a growing number of studies \citep{Zorita1999, Wetterhall2005, Wetterhall2007, Matulla2007, Chardon2014, Dayon2015}.

\citet{Hamill2006} used an analogy based approach on the GFS reforecasts in order to correct systematic errors in the ensemble forecasts of temperature and precipitation. Indeed, the statistical approach showed some bias in estimates of the numerical model, which may have been adjusted by taking into account the intrinsic local climatology from the analogues method. Moreover, the under-dispersion of the ensemble forecast from the numerical model has been corrected using analogues \citep{Hamill2006}. Correction of ensemble forecast under-dispersion by means of the analogue method is also used operationally at EDF.

\citet{Bliefernicht2010} observed that the performance of the analogue method is higher for winter than for summer. The relationship between synoptic predictors and local rainfall is lower in summer, due to convective precipitation that present a higher spatial variability and that depend on other parameters. Variables describing the synoptic circulation are indeed not able to predict the location of thunderstorm cells. This was also observed by \citet{BenDaoud2010}, who set up a specific model for the summer months (June 15 to September 15).

One of the method limitations is the need for a substantial archive of the predictand variable, for example measured precipitations. Without data on several decades, the method is not applicable. Long predictor archives are also required, but this problem is usually solved with reanalysis data, which are not perfect in terms of homogeneity, but can be considered of sufficient quality. Moreover,  reanalysis data are available all around the world, which represent a great potential for the analogue method. Another issue, still about data, concerns the operational forecasting: predictors describing forecasted target situations and the ones from the archive are not fully homogeneous, as they usually don't result from the same model. It is therefore necessary to use robust predictors that depend from the numerical model characteristics at a minimum extent.

Another limitation is the fact that extreme events may be under-represented in the considered sample of analogue dates. Indeed, in a limited weather archive, events with high return periods are not so numerous. Their number is certainly lower than the number of analogues we consider, which can introduce a bias in the forecast. There are however techniques to correct this bias \citep[see][]{Marty2010}.

It is legitimate to raise the question of the relevance of an approach based on archives of past situations in a context of climate change. The first potential issue is a possible change in atmospheric circulation. \citep{Philipp2007} observed certain trends in the synoptic atmospheric circulation over Europe, but with moderate frequencies. Moreover, the basic laws governing the atmosphere behavior will not be transformed \citep{Hewitson1996}. The assumption is that a large part of local climate change will result from changes in intensity, frequency and persistence of synoptic variables, but with other characteristics substantially similar to the present situation \citep{Hewitson1996}. Thus, if the archive of weather situations is long enough, it is reasonable to assume that a large part of future situations is already represented, even those whose frequency will change under different climatic conditions \citep{Wetterhall2005}. In addition, climate change is relatively slow, and through continuous updating of the archive, the analogue method will incorporate new information progressively with no big discontinuity.
